---
title: "Machine Learning Algortithm for Personal Activiy Monitor."
output: html_document
---
###background
####Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement ¨C a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

####Based on the data recorded from several people using Activity Monitors, we will try to construct a macine learning algorithm to predict future outcomes on the test data. I am using Random Forest Modeling, and got over 99% accuracy on my training set of 70% of the total data. All my submitted test cases were succesful.

###Loading and cleaning Data

```{r}
library(caret)
testBulk <- read.csv("pml-testing.csv",na.strings=c("NA",""))
trainBulk <- read.csv("pml-training.csv",na.strings=c("NA",""))
NAs <- apply(trainBulk,2,function(x) {sum(is.na(x))}) 
cleanTrain <- trainBulk[,which(NAs == 0)]
cleanTest <- testBulk[,which(NAs == 0)]
```
###Building data sets for training and cross validation.

####Using 70% for training and 30% for Cross Validation. None generated for testing since that set is already provided.
```{r}
trainIndex <- createDataPartition(y = cleanTrain$classe, p=0.7,list=FALSE)
trainSet <- cleanTrain[trainIndex,]
crossValidationSet <- cleanTrain[-trainIndex,]
# Removing variables that have time, or names in it, also new_window. Columns 1..6
removeIndex <- as.integer(c(1,2,3,4,5,6))
trainSet <- trainSet[,-removeIndex]
testSet <- cleanTest[,-removeIndex]
```

###Training...Training...
```{r}
mytrControl = trainControl(method = "cv", number = 4)
modelFit <- train(trainSet$classe ~.,data = trainSet, method="rf", trControl = mytrControl)
modelFit
```

###Calculation the errors using the Cross Validation Set.
```{r}
predicted <- predict(modelFit, crossValidationSet)
SampleError <- sum(predicted == crossValidationSet$classe)/nrow(crossValidationSet)
```
####So the Out of Sample Error we get is: r SampleError

###Generating data for the prediction vector for the Assigment Submission
```{r}
answers <- predict(modelFit, testSet)
```

